CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-09-11 PaperRMFI2019/samplepaper.tex
update text

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named this 
measurements body intuitive controls (\cite{Allen.2003}), biometric 
measurements (\cite{Sigal.2008}), body dimensions 
(\cite{DBLP:conf/bmvc/ChenRC11}), semantical parameters (\cite{Yang.2014}), 
traditional anthropometric 
measurements 
(\cite{Wuhrer2011}) or only ``shape" as in human shape estimation 
(\cite{Guan.2013}, \cite{Bogo:ECCV:2016}, \cite{Loper.2015}, 
\cite{Dibra.2016a}, \cite{Pishchulin.2017}). In contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements.

@@

In this work we synthesize 3D human meshes using the SMPL
model \cite{Loper.2015}. This model is at its top level a skinned articulated 
model, i.e., 
consists of a 
surface mesh that mimics the skin and a skeleton related to that mesh. It is 
defined by a mean 
template shape represented by a vector of $N$ concatenated vertices 
$\mathbf{\bar{T}}$ in the zero pose, $\vec{\theta}^*$. In order 
to 
synthesize a 
new human mesh one has to deform the provided template mesh by 
setting shape parameters $\vec{\beta}$ and pose parameters $\vec{\theta}$. The 
model provides learned parameters
\begin{equation} \label{eq:smpl_params}
\Phi = \{\mathbf{\bar{T}}, \mathcal{W}, \mathcal{S}, \mathcal{J}, 
\mathcal{P}\}
\end{equation}
As mentioned above $\mathbf{\bar{T}}$ is the mean template shape. The weight 
matrix $\mathcal{W}$ represents how much the rotation of skeleton parts affects 
the vertices. In addition, the matrices $\mathcal{S}$ and $\mathcal{P}$ define 
linear functions that are used to deform $\mathbf{\bar{T}}$ and the matrix 
$\mathcal{J}$ predicts skeleton rest joint locations from vertices in the rest 
pose. We held fix these parameters during the synthesis.

@@

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the pelvis 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves \cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j \in 
\mathbb{R}$ as
\begin{equation}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
	1}^{i = n}s_i
\end{equation}

@@

By comparing neighboring values of the mesh signature $\mathcal{MS}$ as defined 
in equation \ref{eq:mesh_signature} we find several \textit{extrema}. In 
general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms. From a topographic point of view $M_g$ is the highest peak 
	of $\mathcal{MS}$. A horizontal line at height proportion $h$ of the peak 
	intersects 
	it at two points left $p_l$ and right $p_r$.
	\item The local maximum $M_{pc}$ with largest value other than $M_g$ 
	immediately posterior to $p_r$ represents the pelvis circumference.
	\item The local minimum $M_{wc}$ posterior to $p_r$ but prior to $M_{pc}$ 
	represents waist circumference.
	\item The local maximum $M_{cc}$ posterior to $p_r$ but prior to $M_{wc}$ 
	represents 
	chest circumference. If this local maximum does not exist we set $M_{cc}$ 
	to the value of the path $\mathcal{BL}(j_{cc})$ with largest value in the 
	interval $(p_l; M_{wc})$.
\end{enumerate}

>>>>

\keywords{Human Shape Estimation \and CNN \and Another keyword.}
\end{abstract}
%
%
%
\textbf{Motivation} 
Predicting 3D human body measurements from images is crucial in several 
scenarios 
like virtual try-on, animating, ergonomics, computational forensics and 
even health and mortality estimation. Researches had 
named this 
measurements body intuitive controls (\cite{Allen.2003}), biometric 
measurements (\cite{Sigal.2008}), body dimensions 
(\cite{DBLP:conf/bmvc/ChenRC11}, European standard EN 13402), semantical 
parameters (\cite{Yang.2014}), 
traditional anthropometric 
measurements 
(\cite{Wuhrer2011}) or only ``shape" as in human shape estimation 
(\cite{Guan.2013}, \cite{Bogo:ECCV:2016}, \cite{Loper.2015}, 
\cite{Dibra.2016a}, \cite{Pishchulin.2017}). In contrast, we assume a more 
anthropometric approach 
motivated by comprehensive compendiums like Panero and Zelnik, 1979 
\cite{panero1979human}.
Throughout this paper the term human body 
dimensions 
will be used to refer to the above measurements.

@@

In this work we synthesize 3D human meshes using the SMPL
model \cite{Loper.2015}. The generative nature of our approach establishes this 
model as starting point (and not the 3D mesh). Nevertheless, our method is 
flexible enough to begin the pipeline with a 3D mesh. In that case, an SMPL 
model can be fitted to it, using the method of (rh) for example.
This model is at its top level a skinned articulated 
model, i.e., 
consists of a 
surface mesh that mimics the skin and a skeleton related to that mesh. It is 
defined by a mean 
template shape represented by a vector of $N$ concatenated vertices 
$\mathbf{\bar{T}}$ in the zero pose, $\vec{\theta}^*$. In order 
to 
synthesize a 
new human mesh one has to deform the provided template mesh by 
setting shape parameters $\vec{\beta}$ and pose parameters $\vec{\theta}$. The 
model provides learned parameters
\begin{equation} \label{eq:smpl_params}
\Phi = \{\mathbf{\bar{T}}, \mathcal{W}, \mathcal{S}, \mathcal{J}, 
\mathcal{P}\}
\end{equation}
As mentioned above $\mathbf{\bar{T}}$ is the mean template shape. The weight 
matrix $\mathcal{W}$ represents how much the rotation of skeleton parts affects 
the vertices. In addition, the matrices $\mathcal{S}$ and $\mathcal{P}$ define 
linear functions that are used to deform $\mathbf{\bar{T}}$ and the matrix 
$\mathcal{J}$ predicts skeleton rest joint locations from vertices in the rest 
pose. We held fix these parameters during the synthesis.

@@

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. This intuition is compliant 
with prior body dimensions standarised definitions (rf EN 13402 - 1). 
Similarly, the pelvis 
circumference is measured often (rf) around the widest part of hips and 
buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves \cite{book.compu.topo} along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j \in 
\mathbb{R}$ as
\begin{equation}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
	1}^{i = n}s_i
\end{equation}

@@

By comparing neighboring values of the mesh signature $\mathcal{MS}$ as defined 
in equation \ref{eq:mesh_signature} we find several \textit{extrema}. In 
general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. However, there is no consensus in the 
literature regarding the method to calculate chest, waist and hip 
circumference. While Yansel et al., 2020 define waist circumference as the

More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms. From a topographic point of view $M_g$ is the highest peak 
	of $\mathcal{MS}$. A horizontal line at height proportion $h$ of the peak 
	intersects 
	it at two points left $p_l$ and right $p_r$.
	\item The plane $\boldsymbol{\pi}$ intersects the mesh $\mathcal{M}$ at the 
	point where the left hip is located defining a path. The length of this 
	path is $P_{lh}$.	
	\item The local maximum $M_{pc}$ with largest value other than $M_g$ 
	immediately prior to $P_{lh}$ represents the pelvis circumference.	
	\item The local minimum $M_{wc}$ posterior to $p_r$ but prior to $M_{pc}$ 
	represents waist circumference.
	\item The local maximum $M_{cc}$ posterior to $p_r$ but prior to $M_{wc}$ 
	represents 
	chest circumference. If this local maximum does not exist we set $M_{cc}$ 
	to the value of the path $\mathcal{BL}(j_{cc})$ with largest value in the 
	interval $(p_l; M_{wc})$.
\end{enumerate}
